---
title: "2024-2026: How much have I read?"
---

I love to read - I started getting into reading in the fourth grade, and ever since then, I average over 100 books a year. I know briefly that my most read genres are fantasy and romance, but I am curious to see whether those trends hold up when I actually analyze them. I logged all the books I read, and when I read them, from 2024 onwards on a platform called StoryGraph. I exported this data, added some more columns that I felt were important, to get a dataset for me to work with. 


```{r, warning=FALSE}
#| code-fold: true
#| code-summary: "Set-up Code"

#Load Packages 
library(tidyverse)
library(patchwork)
library(grid)
library(cowplot)

#Read in Data
books = read_csv("storygraph.csv")
```

```{r, warning=FALSE}
#| code-fold: true
#| code-summary: "Initial Wrangling Code"

#Remove unnecessary columns 
books = books%>%
  select(-`Date Added`, -`ISBN/UID`, -`Read Count`)

#Rename the columns
books = books%>%
  rename(genre_1 = `Genre 1`, genre_2 = `Genre 2`, pages = `Page Count`, status = `Read Status`, last_date = `Last Date Read`, date_range = `Dates Read`, authors=Authors, title=Title, series = Series)

#Remove whitespace around all the values in all the columns
books = books %>%
  mutate(across(where(is.character), str_trim))

#Keep only books that I read 
books = books %>%
  filter(status == "read")

#Change last date to be a date variable
books = books%>%
  mutate(last_date = as.Date(last_date))

#Remove books read before 2024 (the records are not complete before that) 
books = books %>%
  filter(year(last_date) >= 2024)
```


### Books Read Over Time

I am first interested in seeing how many books I have read over time, which I can also disaggregate by genre. I seem to read a lot during the summer and in the winter, which makes sense, because I am not at college during those times. I also seem to read a lot of romance and fantasy books. I used the read a lot more sports books, but that has decreased a lot, to no sports books in the last few months.  

```{r, warning=FALSE}
#| code-fold: true
#| code-summary: "Data Wrangling Code"

genres = c("Fantasy", "Romance", "Contemporary",
            "Crime", "Mystery", "Sports")

# Create a sequence of every month in the period
all_months = tibble(
  month = seq(from = as.Date("2024-01-01"), to = as.Date("2026-08-01"), by = "month"))

# Create monthly book counts
monthly_books = books %>%
  mutate(month = floor_date(last_date, unit = "month"),
         genre_1 = str_to_lower(genre_1),
         genre_2 = str_to_lower(genre_2)) %>%
  filter(month >= as.Date("2024-01-01"),
         month <= as.Date("2026-08-01")) %>%
  group_by(month) %>%
  summarise(total = n(),
    Fantasy = sum(genre_1 == "fantasy" | genre_2 == "fantasy", na.rm = TRUE),
    Romance = sum(genre_1 == "romance" | genre_2 == "romance", na.rm = TRUE),
    Contemporary = sum(genre_1 == "contemporary" | genre_2 == "contemporary", na.rm = TRUE),
    Crime = sum(genre_1 == "crime" | genre_2 == "crime", na.rm = TRUE),
    Mystery = sum(genre_1 == "mystery" | genre_2 == "mystery", na.rm = TRUE),
    Sports = sum(genre_1 == "sports" | genre_2 == "sports", na.rm = TRUE),
    .groups = "drop") %>%
  right_join(all_months, by = "month") %>%
  mutate(across(-month, ~ replace_na(.x, 0))) %>%
  arrange(month)

monthly_books_long = monthly_books %>%
  pivot_longer(cols = c(Fantasy, Romance, Contemporary, Crime, Mystery, Sports),
    names_to = "genre",
    values_to = "books")

monthly_books_long = monthly_books_long%>%
  select(-total)
```

```{r, warning=FALSE, width = 20}
#| code-fold: true
#| code-summary: "Graphing Code"

#Total Plot 
total_plot = ggplot(monthly_books, aes(x = month, y = total)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b\n%Y") +
  labs(title = "Total Books Read by Month, from January 2024 to August 2026", subtitle = "I read most frequently in the summer; My most read genres are romance and fantasy", x = NULL, y = NULL) +
  theme_classic() +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank(), plot.title = element_text(size = 14))

#By genre Plot
genre_plot = ggplot(monthly_books_long, aes(x = month, y = books, color = genre)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  geom_point(size = 2, alpha = 0.7) +
  scale_x_date(date_breaks = "1 month",
    date_labels = "%b\n%Y") +
  labs(title = "Books Read by Month, Separated by Genre", caption = "Date Created: September 29, 2026 | Author: Tanisha | Data: StoryGraph", y = NULL, color = NULL, x= NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 7),
      legend.position = "bottom",
    legend.title = element_blank(),
    legend.direction = "horizontal", plot.title = element_text(size = 14))+
  guides(color = guide_legend(nrow = 1))

#Combine plots 
combined_plot = plot_grid(
  total_plot,
  genre_plot,
  ncol = 1,
  align = "v",
  axis = "lr",
  rel_heights = c(1, 2))

final_plot <- plot_grid(
  ggdraw() +
    draw_label(
      "Number of Books Read",
      angle = 90,
      x = 0.5,
      y = 0.65, size = 10),
  combined_plot,
  ncol = 2,
  rel_widths = c(0.05, 1))

final_plot
```

### Genres by pages

I read mostly romance books, followed by fantasy. 


```{r, warning=FALSE, width = 20}
#| code-fold: true
#| code-summary: "Genre Bar Plot Code"

#Number of books by genre
genre_bar = books %>%
  summarise(
    Fantasy = sum(genre_1 == "Fantasy" | genre_2 == "Fantasy", na.rm = TRUE),
    Romance = sum(genre_1 == "Romance" | genre_2 == "Romance", na.rm = TRUE),
    Contemporary = sum(genre_1 == "Contemporary" | genre_2 == "Contemporary", na.rm = TRUE),
    Crime = sum(genre_1 == "Crime" | genre_2 == "Crime", na.rm = TRUE),
    Mystery = sum(genre_1 == "Mystery" | genre_2 == "Mystery", na.rm = TRUE),
    Sports = sum(genre_1 == "Sports" | genre_2 == "Sports", na.rm = TRUE))%>%
  pivot_longer(cols = everything(), names_to = "genre", values_to = "books")

#Number of pages by genre
genre_pages = books %>%
  summarise(Fantasy = sum(pages[genre_1 == "Fantasy" | genre_2 == "Fantasy"], na.rm = TRUE),
            Romance = sum(pages[genre_1 == "Romance" | genre_2 == "Romance"], na.rm = TRUE),
            Contemporary = sum(pages[genre_1 == "Contemporary" | genre_2 == "Contemporary"], na.rm = TRUE),
            Crime = sum(pages[genre_1 == "Crime" | genre_2 == "Crime"], na.rm = TRUE),
            Mystery = sum(pages[genre_1 == "Mystery" | genre_2 == "Mystery"], na.rm = TRUE),
            Sports = sum(pages[genre_1 == "Sports" | genre_2 == "Sports"], na.rm = TRUE)) %>%
  pivot_longer(cols = everything(), names_to = "genre", values_to = "pages")


#merge on genre
genre_data = genre_bar %>%
  left_join(genre_pages, by = "genre")

genre_bar_long = genre_data %>%
  select(genre, books, pages) %>%
  pivot_longer(cols = c(books, pages), names_to = "measure", values_to = "value")

# Find a scaling factor between pages and books
scale_factor <- max(genre_data$pages) / max(genre_data$books)

ggplot(genre_data, aes(x = genre)) +
  
  # Books
  geom_col(
    aes(y = books, fill = "Books"),
    position = position_nudge(x = -0.2),
    width = 0.35
  ) +
  
  # Pages, scaled to book axis
  geom_col(
    aes(y = pages / scale_factor, fill = "Pages"),
    position = position_nudge(x = 0.2),
    width = 0.35
  ) +
  
  scale_y_continuous(
    name = "Number of Books",
    sec.axis = sec_axis(
      ~ . * scale_factor,
      name = "Number of Pages"
    )
  ) +
  
  labs(
    title = "Books and Pages Read by Genre", subtitle = "I read mostly romance books, followed by fantasy", caption = "Date Created: September 29, 2026 | Author: Tanisha | Data: StoryGraph",
    x = NULL,
    fill = NULL
  ) +
  
  theme_classic()
```



